% arara: xelatex
\documentclass[12pt]{article}

\usepackage{physics}


\usepackage{tikz} % картинки в tikz
\usepackage{microtype} % свешивание пунктуации

\usepackage{array} % для столбцов фиксированной ширины

\usepackage{indentfirst} % отступ в первом параграфе

\usepackage{sectsty} % для центрирования названий частей
\allsectionsfont{\centering}

\usepackage{amsmath, amsfonts, amssymb} % куча стандартных математических плюшек

\usepackage{comment}

\usepackage[top=2cm, left=1.2cm, right=1.2cm, bottom=2cm]{geometry} % размер текста на странице

\usepackage{lastpage} % чтобы узнать номер последней страницы

\usepackage{enumitem} % дополнительные плюшки для списков
%  например \begin{enumerate}[resume] позволяет продолжить нумерацию в новом списке
\usepackage{caption}

\usepackage{url} % to use \url{link to web}

\usepackage{fancyhdr} % весёлые колонтитулы
\pagestyle{fancy}
\lhead{Statistics 4 Market Research}
\chead{}
\rhead{HA}
\lfoot{2022-2023}
\cfoot{}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}



\usepackage{todonotes} % для вставки в документ заметок о том, что осталось сделать
% \todo{Здесь надо коэффициенты исправить}
% \missingfigure{Здесь будет Последний день Помпеи}
% \listoftodos - печатает все поставленные \todo'шки


% более красивые таблицы
\usepackage{booktabs}
% заповеди из докупентации:
% 1. Не используйте вертикальные линни
% 2. Не используйте двойные линии
% 3. Единицы измерения - в шапку таблицы
% 4. Не сокращайте .1 вместо 0.1
% 5. Повторяющееся значение повторяйте, а не говорите "то же"



\usepackage{fontspec}
\usepackage{polyglossia}

\setmainlanguage{english}
\setotherlanguages{english}

% download "Linux Libertine" fonts:
% http://www.linuxlibertine.org/index.php?id=91&L=1
\setmainfont{Linux Libertine O} % or Helvetica, Arial, Cambria
% why do we need \newfontfamily:
% http://tex.stackexchange.com/questions/91507/
\newfontfamily{\cyrillicfonttt}{Linux Libertine O}

%\AddEnumerateCounter{\asbuk}{\russian@alph}{щ} % для списков с русскими буквами
%\setlist[enumerate, 2]{label=\asbuk*),ref=\asbuk*}

%% эконометрические сокращения
\DeclareMathOperator{\Cov}{\mathbb{C}ov}
\DeclareMathOperator{\Corr}{\mathbb{C}orr}
\DeclareMathOperator{\Var}{\mathbb{V}ar}

\let\P\relax
\DeclareMathOperator{\P}{\mathbb{P}}
\DeclareMathOperator{\plim}{\mathrm{plim}}

\DeclareMathOperator{\E}{\mathbb{E}}
% \DeclareMathOperator{\tr}{trace}
\DeclareMathOperator{\card}{card}
\DeclareMathOperator{\pCorr}{\mathrm{p}\mathbb{C}\mathrm{orr}}


\newcommand \hb{\hat{\beta}}
\newcommand \hs{\hat{\sigma}}
\newcommand \htheta{\hat{\theta}}
\newcommand \s{\sigma}
\newcommand \hy{\hat{y}}
\newcommand \hY{\hat{Y}}
\newcommand \e{\varepsilon}
\newcommand \he{\hat{\e}}
\newcommand \z{z}
\newcommand \hVar{\widehat{\Var}}
\newcommand \hCorr{\widehat{\Corr}}
\newcommand \hCov{\widehat{\Cov}}
\newcommand \cN{\mathcal{N}}
\newcommand \RR{\mathbb{R}}
\newcommand \NN{\mathbb{N}}
\newcommand{\cF}{\mathcal{F}}
\newcommand{\cH}{\mathcal{H}}


\begin{document}

\section*{Home Assignment 1}
\begin{enumerate}

  \item Consider naive bootstrap for a sample of $n$ observations. We know that $y_i$ are iid uniform on $[0;10]$.
  \begin{enumerate}
    \item What is the probability that original $y_5$ will enter any particular bootstrap sample exactly $3$ times?
    \item What is the limit of probability you found in (a) when $n\to\infty$?
    \item What is the limit of probability $\P(\max y_i > \max y_i^*)$ when $n\to\infty$?
    \item Find $\E(y_1^*)$, $\Var(y_1^*)$, $\Cov(y_1^*, y_2^*)$.
  \end{enumerate}

  \item Winnie-the-Pooh has collected 15 honey samples from good (G) and bad (B) bees. 
  Winnie has ordered the honey quality from lower to higher on the real line.
  He has forgot the exact values, but here is the ordering: BBBGB-BGGBG-GGBGG.

  \begin{enumerate}
    \item Please, help Winnie and test the null-hypothesis that the honey quality is the same 
    against alternative that good bees produce better honey at 5\% significance level.
    \item Using computer find the exact P-value of this ordering.
  \end{enumerate}
  
  \item What should be the minimal total sample size $n$ for the following testing problem?
  
  The goal is to have first type error $\alpha \leq 0.01$, second type error $\beta \leq 0.1$,
  minimal detectable difference in probabilities should be $H_A: p_B - p_A \geq 0.2$, 60\% of 
  total sample should go into test group $B$ and 40\% — into control group $A$.
  
  \item Winnie-the-Pooh simulteneously tests four hypothesis. 
  Actually three null hypothesis are true and one is false (the corresponding p-value is uniform on $[0;0.5]$),
  but Winnie does not know this. 
  

  \begin{enumerate}
    \item Winnie uses Bonferonni correction with target bound $0.05$ for $FWER$. 
    What is the actual $FWER$?
    \item Winnie uses Holm-Bonferonni correction with target bound $0.05$ for $FWER$. 
    What is the actual $FWER$?
  \end{enumerate}
  \item The sample correlation matrix of three variables $(a, b, c)$ is 
  \[
C = \begin{pmatrix}
  1 & 0.7 & 0  \\
  0.7 & 1 & 0 \\
  0 & 0 & 1 \\
\end{pmatrix}.
  \]
  The number of observations is equal to $n=101$.

  \begin{enumerate}
    \item State the weights of $a$, $b$ and $c$ in the first principal component. 
    \item All you know about observation number 7 that the first principal component is equal to $0.1$ and the second — to $0.2$.
    Recover approximate values of $a_7$, $b_7$ and $c_7$.
    \item How much variance is explained by the first two components?
  \end{enumerate}


\end{enumerate}

Deadline: \textbf{2022-12-06, 21:00}. 


\end{document}

