% arara: xelatex: {shell: yes}
%% arara: biber
%% arara: xelatex: {shell: yes}
%% arara: xelatex: {shell: yes}
\documentclass[12pt]{article}

\usepackage{etoolbox} % для условия if-else
\newtoggle{excerpt} % помечаем, что это отрывок, а далее в тексте может использовать
\toggletrue{excerpt}
% команду \iftoggle{excerpt}{yes}{no}

\usepackage{tikz} % картинки в tikz
\usepackage{microtype} % свешивание пунктуации

\usepackage{array} % для столбцов фиксированной ширины

\usepackage{indentfirst} % отступ в первом параграфе

\usepackage{sectsty} % для центрирования названий частей
\allsectionsfont{\centering}

\usepackage{amsmath} % куча стандартных математических плюшек

\usepackage{comment}
\usepackage{amsfonts}

\usepackage[top=2cm, left=1.2cm, right=1.2cm, bottom=2cm]{geometry} % размер текста на странице

\usepackage{lastpage} % чтобы узнать номер последней страницы

\usepackage{enumitem} % дополнительные плюшки для списков
%  например \begin{enumerate}[resume] позволяет продолжить нумерацию в новом списке
\usepackage{caption}


\usepackage{fancyhdr} % весёлые колонтитулы
\pagestyle{fancy}
\lhead{Statistics for Market Research}
\chead{}
\rhead{Retake, 2023-01-23}
\lfoot{}
\cfoot{}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}



\usepackage{todonotes} % для вставки в документ заметок о том, что осталось сделать
% \todo{Здесь надо коэффициенты исправить}
% \missingfigure{Здесь будет Последний день Помпеи}
% \listoftodos --- печатает все поставленные \todo'шки


% более красивые таблицы
\usepackage{booktabs}
% заповеди из докупентации:
% 1. Не используйте вертикальные линни
% 2. Не используйте двойные линии
% 3. Единицы измерения - в шапку таблицы
% 4. Не сокращайте .1 вместо 0.1
% 5. Повторяющееся значение повторяйте, а не говорите "то же"

\usepackage{multicol}


\usepackage{fontspec}
\usepackage{polyglossia}

\setmainlanguage{russian}
\setotherlanguages{english}

% download "Linux Libertine" fonts:
% http://www.linuxlibertine.org/index.php?id=91&L=1
\setmainfont{Linux Libertine O} % or Helvetica, Arial, Cambria
% why do we need \newfontfamily:
% http://tex.stackexchange.com/questions/91507/
\newfontfamily{\cyrillicfonttt}{Linux Libertine O}

\AddEnumerateCounter{\asbuk}{\russian@alph}{щ} % для списков с русскими буквами
\setlist[enumerate, 2]{label=\alph*),ref=\alph*} % \alph* \asbuk* \arabic*

%% эконометрические сокращения
\DeclareMathOperator{\Cov}{Cov}
\DeclareMathOperator{\Corr}{Corr}
\DeclareMathOperator{\Var}{Var}
\DeclareMathOperator{\E}{E}
\newcommand \hb{\hat{\beta}}
\newcommand \hs{\hat{\sigma}}
\newcommand \htheta{\hat{\theta}}
\newcommand \s{\sigma}
\newcommand \hy{\hat{y}}
\newcommand \hY{\hat{Y}}
\newcommand \e{\varepsilon}
\newcommand \he{\hat{\e}}
\newcommand \z{z}
\newcommand \hVar{\widehat{\Var}}
\newcommand \hCorr{\widehat{\Corr}}
\newcommand \hCov{\widehat{\Cov}}
\newcommand \cN{\mathcal{N}}
\let\P\relax
\DeclareMathOperator{\P}{\mathbb{P}}

\DeclareMathOperator{\plim}{plim}



\usepackage{color,url,amsthm,amssymb,longtable,eurosym}
\newenvironment{question}{\item }{}
% \newenvironment{solution}{}{}
\excludecomment{solution}
\newenvironment{answerlist}{\begin{multicols}{3}\begin{enumerate}}{\end{enumerate}\end{multicols}}
\newcommand{\setzerocols}{
  \renewenvironment{answerlist}{\begin{enumerate}}{\end{enumerate}}
}
\newcommand{\setncols}[1]{
  \renewenvironment{answerlist}{\begin{multicols}{#1}\begin{enumerate}}{\end{enumerate}\end{multicols}}
}


\newcommand{\answerbox}{\raisebox{3mm}{%
    \fbox{%
          \begin{minipage}[t]{2mm}%
              \hspace*{2mm}%
              \vspace*{0.3cm}
          \end{minipage}
         }
    }
}

\newcommand{\answerline}{
    \answerbox\hspace*{-1mm}a \hspace*{3mm}
    \answerbox\hspace*{-1mm}b \hspace*{3mm}
    \answerbox\hspace*{-1mm}c \hspace*{3mm}
    \answerbox\hspace*{-1mm}d \hspace*{3mm}
    \answerbox\hspace*{-1mm}e \hspace*{3mm}
    \answerbox\hspace*{-1mm}f \hspace*{3mm}
}

\newcommand{\fiobox}{
\fbox{
  \begin{minipage}{42em}
    Имя, фамилия и номер группы:\vspace*{3ex}\par
    \noindent\dotfill\vspace{2mm}
  \end{minipage}
}
}

% делаем короче интервал в списках
\setlength{\itemsep}{0pt}
\setlength{\parskip}{0pt}
\setlength{\parsep}{0pt}

\begin{document}

Part I. Only numeric answer is checked, you can attempt all questions but the maximal grade is 4 points. 
	\begin{enumerate}
	\item I have a sample \(X_1, \ldots, X_{90}\).
	I generate one naive bootstrap sample \(X^*_1, \ldots, X^*_{90}\).
	
	Let \(N\) be the number of times the first observation will be copied in the bootstrap sample.
	
	
	What is the probability that the first observation will be present in the bootstrap sample more than once?
	
		
	
	\item We have data of an AB experiment: \(\bar X_a = 5.4\), \(\bar X_b = 6\),
		\(n_a = 18\), \(n_b = 15\), \(\sum (X_i^a - \bar X_a)^2 = 890\),
		\(\sum (X_i^b - \bar X_b)^2 = 800\).

		Calculate the estimate of variance of \(\bar X_a - \bar X_b\) for the Welch test.

		\item I have five results of two runners A and B for the 5 km race:

		16:49 (B), 21:17 (A), 18:30 (B), 6:18 (B), 20:16 (A), 15:39 (B).
		
		Calculate Mann-Whitney statistic \(U_A\) that tests the null-hypothesis of equal distributions of time.
		(The statistic \(U_A\) should positively depend on the ranks of the runner A).
		\item I have 30 hypothesis with independent statistics. The null hypothesis for all 30
		cases is actually true, but I don't know this.
		
		I calculate all p-values.
		If the 4 lowest p-value are simultaneously lower than 0.01 I wrongly conclude that not all \(H_0\) are true. Otherwise I correctly conclude that all \(H_0\) are true.
		
		What is the probability that I will get the correct conclusion?
		
		\item My target variable is binary and I wish minimal detectable effect equal to 0.04, probability of I-error not greater than 0.01, probability of II-error not greater than 0.3. The control and experimental group
		are of the same size equal to \(n\).
		
		Which minimal value of \(n\) is sufficient in the worst case?
		\item Vasiliy loves to eat shaurma. He has \(6\) local shaurma dealers. Vasiliy bought \(5\) shaurmas
from each dealer and measured their weight. He would like to test the hypothesis that mean weight is the
same for all dealers.

Total sum of squares is \(700\), between sum of squares is \(200\)
Calculate the F -statistic to test the hypothesis.
		
		
	\end{enumerate}
	\newpage

	Part II. Full solution is required, you can attempt all questions but the maximal grade is 6 points. 
	\begin{enumerate}[resume]
			\item Let random variables  be iid uniform $U[0;1]$.
			Consider the original sample $Y_1$, \ldots, $Y_n$ and naive bootstrap sample $Y_1^*$, \ldots, $Y_n^*$.
			
			Let $N_1$ be the number of copies of $Y_1$ in the bootstrap sample. 

			Find the distribution of $N_1$ when $n\to \infty$. 

			\item I have invented the GCA, \textit{garbage component analysis}. 
			It is completely similar to PCA with one exception: you \textit{minimize} the variance of components. 
			First you find the component with minimal sample variance. Than you find the second component that is 
			orthogonal to the first one with minimal variance. And so on. 

			How my new garbage components are related to classic principal components?


			\item Winnie-the-Pooh simulteneously tests $h$ null hypothesis using independent samples. 
			All the null hypothesis are true but Winnie does not know it. 
			
			\begin{enumerate}
				\item What is the expected value of the second lowest P-value?
				\item What is the variance of the number of wrongly regected hypothesis if Winnie rejects all the hypothesis with P-value less $0.05$?
			\end{enumerate}
		
		\item There are three continuosly distributed samples of the same size $n$, $X_1$, \ldots, $X_n$, $Y_1$, \ldots, $Y_n$,
		$Z_1$, \ldots, $Z_n$. Imagine that the null hypothesis that all samples have the same distribution is true. 
		
		Consider the random variable $R_X$ — the sum of ranks of the $X$ sample in the pooled sample. 
		
		Find the correlation $\Corr(R_X, R_Y)$.
		\end{enumerate}
		
	

\end{document}

